---
title: "Knife-Edge Scan"
output: html_notebook
---
```{r}
library(smoother)
source("ElementalImage.R")
source("rsum.R")


```

Knife-edge scan to determine the spot size of the µXRF instrument in the Fittschen Group at the Washington State University, Pullman, WA

##Fe-Knife Edge (Razor Blade)

We collect several scans in a row and use PyMCA to batch-fit the measured spectra and determine the counts for specific elements/lines

```{r}
readPyMCAData("data/B0216-1325-000001_to_000480.dat")
scanData <- PyMCAData$Fe.Ka

plot(scanData, type="l",ylab="counts")
title(quote(Fe ~ K*alpha))
```
```{r}
numScans <-6
samplesPerScan <- length(scanData)/numScans
stepWidth <- 2

```

As we can see there are `r numScans` spectra and `r samplesPerScan` samples per scan. The step size was `r stepWidth` µm.

```{r}
knifeList<-split(scanData,rep(1:numScans, each = samplesPerScan))
plot(NULL,xlim = c(0,samplesPerScan*stepWidth), ylim=c(min(scanData),max(scanData)), xlab=quote(position*group("[",mu*m,"]")), ylab="counts",type="n")
title(quote(Fe*K*alpha))
res <- mapply(lines, y=knifeList,MoreArgs = list(x=seq(samplesPerScan)*stepWidth))

```

For the existing functions we need the rising edge of the knife-edge, so we reverse the data. The first scan is taken too early after turning on the tube, so we will remove this scan. 


```{r, message=FALSE}
knifeList<-split(scanData[length(scanData):(1+samplesPerScan)],rep(1:(numScans-1), each = samplesPerScan))
plot(NULL,xlim = c(0,samplesPerScan*stepWidth), ylim=c(min(scanData),max(scanData)), xlab=quote(position*group("[",mu*m,"]")), ylab="counts",type="n")

res <- mapply(lines, y=knifeList,MoreArgs = list(x=seq(samplesPerScan)*stepWidth))
res <- title(quote(Fe*K*alpha))

```
Better... But the z-stage is too incaurate, so we will shift the scans for better alignment. The scns are aligned at the half hight of the y signal. 

```{r fig.width=4, fig.height=4}
par(mar=c(5,4,4,5)+.1)
plotCentered <- function(scanList, width=stepWidth, window=0.15) {
  ycenter <- max(unlist(scanList))/2
  xoffset<-mapply(function(s) {uniroot(approxfun(s-ycenter),lower=1,upper=length(s))$root},scanList)
  plot(NULL,xlim = c(-max(xoffset),samplesPerScan-min(xoffset))*width,type="n",ylim=c(0,2*ycenter), xlab=quote(group("[",mu * m,"]")),ylab="Counts")
  res <- title(quote(Fe*K*alpha* " centered"))
  res <- mapply(function(s,x){lines(x=(seq(samplesPerScan)-x)*width,y=s)},scanList,xoffset)
}

plotCentered(knifeList,stepWidth,10)

```

Looks good. Do we need some statistic verification? Standard deviation,..?



## Beam Width Fe K$\alpha$

The calculation of the FWHM involves numerical differentiation which magnifies noise. To get dependable results we need to filter/smooth the input values. We use a gausian filter, with a parameterized window size. A wider window means moother curves but also a numerical wider spotbsize.  

```{r}
knifedges <- mapply(function(window) {r <- knifeEdge(knifeList[[1]],width=2,window=window); title(paste("Window=",window)); r}, window=c(3,4,5,6,7,8,9,10,12,14,18,22))

```

Standard deviation for 5 measurements with window size=11
```{r}
res <- mapply(knifeEdge,knifeList,MoreArgs = list(width=2, window=11, plotit=FALSE))
knifeEdgeFWHM <- unlist(res["FWHM",])

```

The mean beam size is `r mean(knifeEdgeFWHM)`
The Standard deviation is `r sd(knifeEdgeFWHM)` 
equals `r sd(knifeEdgeFWHM)/mean(knifeEdgeFWHM)*100` %

# Knife Edge Scan Tungsten

Calculated beam width for window sizes 3-14

```{r}
readPyMCAData("data/B0221-1426-000001_to_000050.dat")
scanData <- PyMCAData$W.L3

knifedges <- mapply(knifeEdge, window=c(3:14),MoreArgs = list(scan=rev(scanData),width=2,plotit=FALSE))
library(knitr)
kable(data.frame(width=c(3:14),window=unlist(knifedges["FWHM",1:12])))
```
The selected window width is 9 (due to constraints from the publication ;-)

```{r fig.width=4, fig.height=4}
par(mar=c(5,4,4,5)+.1)
res <- knifeEdge(rev(scanData),2,9,plotit = FALSE)
plot(x=seq_along(scanData)*stepWidth,y=rev(scanData),xlab = quote("distance "*group("[",mu*m,"]")),ylab = "counts")
lines(x=seq_along(scanData)*stepWidth,y=res$smoothSpec)
par(new=TRUE,lty=2)
plot(x=seq_along(res$FWHM.plot)*stepWidth,y=res$FWHM.plot,type="l",xaxt="n",yaxt="n",xlab="",ylab="")
axis(4)
mtext(quote(group("[",counts/mu*m,"]")),side=4,line=3)
par(lty=1)
lines(res$FWHM.coord)
text(res$FWHM.coord$x[2],res$FWHM.coord$y[2],bquote("  FWHM = "*.(round(res$FWHM,1)) ~ mu*m),adj=c(0,0))


```

